---
title: "Data analysis project"
author: "Jess Qualley and Rebecca Hansen"
date: "March 27, 2020"
output: word_document
---

commit test by Jess.

##Clear code (**delete before submitting***)
```{r}
rm(list = ls(all = TRUE))
```

#Preparation
##Loading and citing packages
```{r, echo = TRUE}
#Load faraway package
library(faraway)

#Load plyr package
library(plyr)

#Load tidyverse package
library(tidyverse)

#Cite packages
citation("faraway")
citation("plyr")
citation("tidyverse")
```

##Read in and organize data
```{r, echo = TRUE}
#Read from csv file
herring <- read.csv("HerringOtolithDatabase 15 Jan 2020_JQ.csv", header=TRUE, fileEncoding="UTF-8-BOM")

#Create a data frame
herring_data <- data.frame(sal.id = herring$FishCode,
                           oto.width = herring$AverageWidth,
                           digestion = herring$Digestion,
                           sal.sp = herring$SalmonSpecies,
                           sal.sex = herring$SalmonSex,
                           sal.length = as.numeric(herring$SalmonLength),
                           sal.weight = herring$SalmonWeight,
                           coll.year = herring$CollectionYear,
                           coll.doy =   herring$CollectionDayofYear,
                           lat = herring$Latitude,
                           long = herring$Longitude,
                           stat.area = as.factor(herring$StatArea))

#figure out which stat areas to remove
levels(as.factor(herring$StatArea))

#filter for 2018 data, chinook salmon only, omit northern BC areas (Haid Gwaii) and west coast Vancouver Island
herring_data <- herring_data %>% filter(herring_data$sal.sp == "ch" & herring_data$coll.year == "2018" & herring_data$stat.area != "1" & herring_data$stat.area != "101" & herring_data$stat.area != "125" & herring_data$stat.area != "23" & herring_data$stat.area != "25") %>% na.omit()
```

##Checking our work (***delete before submitting***)
```{r, echo = TRUE}
#plot lat/long as continuous variable
plot(herring_data$long, herring_data$lat, xlim = c(-122, -128), ylim = c(48, 50.5))
```

##Example 1 (***delete before submitting)
```{r, echo = TRUE}
#add column for number of rows
n <- nrow(herring_data)
herring_data$row <- 1:n

par(mfrow = c(3, 2))

#otolith width
with(herring_data, plot(oto.width, row, las = 1, type = "n", xlab = "Otolith width [mm]", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(oto.width, row, las = 1, type = "p", pch = 19, cex = 0.5))

#actual salmon length - all salmon that ate herring - note that some numbers are ranges and have weird symbols...
#where lengths were not recorded, they were calculated using length-weight relationship
with(herring_data, plot(sal.length, row, las = 1, type = "n", xlab = "Actual Salmon Length [mm]", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(sal.length, row, las = 1, type = "p", pch = 19, cex = 0.5))

#actual salmon weight - raw salmon weights no back calculations
with(herring_data, plot(sal.weight, row, las = 1, type = "n", xlab = "Actual Salmon Weight [lbs]", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(sal.weight, row, las = 1, type = "p", pch = 19, cex = 0.5))

#day of year
with(herring_data, plot(coll.doy, row, las = 1, type = "n", xlab = "Day of Year in 2018", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(coll.doy, row, las = 1, type = "p", pch = 19, cex = 0.5))

#lat
with(herring_data, plot(lat, row, las = 1, type = "n", xlab = "Latitude", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(lat, row, las = 1, type = "p", pch = 19, cex = 0.5))

#long
with(herring_data, plot(long, row, las = 1, type = "n", xlab = "Longitude", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(long, row, las = 1, type = "p", pch = 19, cex = 0.5))
```


##Example 2 (***delete before submitting)
```{r, echo = TRUE}
#create data frame for pairwise scatter with only continuous explanatory variables - salmon lenght, weight, collection day of year, latitude and longitude
herring_collinearity <- data.frame("sal.length" = herring_data$sal.length, "sal.weight" = herring_data$sal.weight, "coll.doy" = herring_data$coll.doy, "lat" = herring_data$lat, "long" = herring_data$long)

#create pairwise scatterplot with salmon length, weight and day of year
plot(herring_collinearity[1:5], cex = 0.5, pch = 19, col = rgb(0, 0, 0, 0.5), 
                   labels = c("Chinook\nLength [mm]", "Chinook\nWeight [lbs]", "Collection\nDay in 2018", "Latitude", "Longitude"))

#pairwise correlation coefficients for multiple variables
print(cor(na.omit(herring_data[ ,c("sal.length", "sal.weight", "coll.doy", "lat", "long")])), digits = 2)

#variance inflation factors
print(vif(na.omit(herring_data[ ,c("sal.length", "sal.weight", "coll.doy", "lat", "long")])))
```

##Example 3 (***delete before submitting)
```{r, echo = TRUE}
#Make a table of salmon sex versus year
with(herring_data, table(sal.sex, coll.year))

#Make a table of salmon sex versus digestion
with(herring_data, table(sal.sex, digestion))

#Make a table of salmon year versus digestion
with(herring_data, table(coll.year, digestion))
```

##Example 4 (***delete before submitting)
```{r, echo = TRUE}
#Make a table of 0 values and NAs for oto.width
table(herring_data$oto.width > 0, useNA = "always")

#Make a table of 0 values and NAs for sal.length
table(herring_data$sal.length > 0, useNA = "always")

#Make a table of 0 values and NAs for sal.weight
table(herring_data$sal.weight > 0, useNA = "always")

#Make a table of 0 values and NAs for lat
table(herring_data$lat > 0, useNA = "always")

#Make a table of 0 values and NAs for long
table(herring_data$long > 0, useNA = "always")

#Proportion of measured oto-width values that are zero:
sum(herring_data$oto.width == 0, na.rm = TRUE) / length(na.omit(herring_data$oto.width))

#Proportion of measured sal.length values that are zero:
sum(herring_data$sal.length == 0, na.rm = TRUE) / length(na.omit(herring_data$sal.length))

#Proportion of measured sal.weight values that are zero:
sum(herring_data$sal.weight == 0, na.rm = TRUE) / length(na.omit(herring_data$sal.weight))

#Proportion of measured lat values that are zero:
sum(herring_data$lat == 0, na.rm = TRUE) / length(na.omit(herring_data$lat))

#Proportion of measured long values that are zero:
sum(herring_data$long == 0, na.rm = TRUE) / length(na.omit(herring_data$long))

```

##Cleaning data
```{r, echo = TRUE}
#filter only chinook and data from 2018 with na removed
herring_data_2018 <- herring_data %>% filter(herring_data$sal.sp == "ch" & herring_data$coll.year == "2018") %>% na.omit()
```

#Section 1: Introduction, Question, Goals and Hypotheses (400 words)
##A)
Introduce relevant concepts: 
- State of salmon in BC
- Salmon diet
- Herring as a prey item for salmon
- Gape size theory? Gape size and body size? Prey selection strategies?

How the dataset was collected:
- Salmon stomachs donated to Juanes lab by anglers
- Salmon stomach contents sorted and classfied based on digestion
- Otoliths collected and measured

 Background needed to understand the dataset:
- Otolith width as the best proxy for herring length

At least two scientiﬁc papers that are related to your topic:

##B)
Research question:
- Relationship between salmon size and size of their herring prey

##C)
Goal of research: 
- Inference, we want to test if the size of the salmon determines the size of its herring prey, 

##D)
Hypothesis:
-  hypothesis for how you expect the response variable to change with each predictor variable and interactions: linear or non-linear and what shape? positive or negative? interaction terms? Why?
OR list of alternative hypotheses, in words or as model formulae that correspond to diﬀerent scenarios related to your question
- any null hypotheses

##E)
Model selection:
- Why or why not use it?
- What type to use?

#Section 2: The Response Variable (150 words)
##A)
```{r, echo = TRUE}
#Add column for number of rows
n <- nrow(herring_data)
herring_data$row <- 1:n

#Dotchart of  otolith width (response variable)
with(herring_data, plot(oto.width, row, las = 1, type = "n", xlab = "Otolith width [mm]", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(oto.width, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Histogram of response variable with rugmarks
hist(herring_data$oto.width, xlim = c(0.3, 2.8), breaks = seq(0.2, 2.8, by = 0.05), col = grey(0.9), main = "Histogram of herring otolith width", xlab = "Herring otolith width")
rug(herring_data$oto.width)
```

##B) 
Qualitative description of response variable:
- Continuous 
- Bounded by 0, can't take negative values

##C)
Potential issues with response variable
- Outliers
- Missing values
- Zeroes
- Non-independence from repeated measurements within experimental units
- Non-idependence from temporal autocorrelation
- Non-independence from spatial autocorrelation

##D)
Candidate distributions for variable
- Family of probability distributions we're likely to use

#Section 3: The Explanatory Variables (150 words)
##A)
###Continuous explanatory variables
```{r, echo = TRUE}
#Multipanel of dotcharts for continuous explanatory variables
par(mfrow = c(3, 2))

#Salmon length
with(herring_data, plot(sal.length, row, las = 1, type = "n", xlab = "Actual Salmon Length [mm]", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(sal.length, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Salmon weight
with(herring_data, plot(sal.weight, row, las = 1, type = "n", xlab = "Actual Salmon Weight [lbs]", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(sal.weight, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Collection date of year
with(herring_data, plot(coll.doy, row, las = 1, type = "n", xlab = "Day of Year in 2018", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(coll.doy, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Latitude
with(herring_data, plot(lat, row, las = 1, type = "n", xlab = "Latitude", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(lat, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Longitude
with(herring_data, plot(long, row, las = 1, type = "n", xlab = "Longitude", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(long, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Multipanel of histograms for continuous explanatory variables
par(mfrow = c(3, 2))

#Salmon length
hist(herring_data$sal.length, col = grey(0.9), main = "Histogram of salmon length", xlab = "Salmon length")
rug(herring_data$sal.length)

#Salmon weight
hist(herring_data$sal.weight, col = grey(0.9), main = "Histogram of salmon weight", xlab = "Salmon weight")
rug(herring_data$sal.weight)

#Collection date of year
hist(herring_data$coll.doy, col = grey(0.9), main = "Histogram of date of year", xlab = "Collection date of year")
rug(herring_data$coll.doy)

#Latitude
hist(herring_data$lat, col = grey(0.9), main = "Histogram of latitude", xlab = "Latitude")
rug(herring_data$lat)

#Longitude
hist(herring_data$long, col = grey(0.9), main = "Histogram of longitude", xlab = "Longitude")
rug(herring_data$long)

```

###Categorical explanatory variables
```{r, echo = TRUE}
#Multipanel of dotcharts for categorical explanatory variables
par(mfrow = c(2, 2))

#Salmon id
with(herring_data, plot(sal.id, row, las = 1, type = "n", xlab = "Salmon ID", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(sal.id, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Digestion
with(herring_data, plot(digestion, row, las = 1, type = "n", xlab = "Digestion state", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(digestion, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Salmon species
with(herring_data, plot(sal.sp, row, las = 1, type = "n", xlab = "Salmon species", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(sal.sp, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Salmon sex
with(herring_data, plot(sal.sex, row, las = 1, type = "n", xlab = "Salmon sex", ylab = "Row Number"))
abline (h = seq(1, n, by = 5), col = "lightgrey")
with(herring_data, points(sal.sex, row, las = 1, type = "p", pch = 19, cex = 0.5))

#Tables for categorical explanatory variables
#Make a table of salmon sex versus year
with(herring_data, table(sal.sex, coll.year))

#Make a table of salmon sex versus digestion
with(herring_data, table(sal.sex, digestion))

#Make a table of salmon year versus digestion
with(herring_data, table(coll.year, digestion))

```

##B)
- Outliers
- Transformations why or why not?

#Section 4: Collinearity, Balance, and Variance Inﬂation Factors (200 words)
##A)
```{r, echo = TRUE}
#create data frame for pairwise scatter with only continuous explanatory variables - salmon lenght, weight, collection day of year, latitude and longitude
herring_collinearity <- data.frame("sal.length" = herring_data$sal.length, "sal.weight" = herring_data$sal.weight, "coll.doy" = herring_data$coll.doy, "lat" = herring_data$lat, "long" = herring_data$long)

#Pairwise scatterplot of explanatory variables (salmon length, weight, day of year, latitude and longitude)
plot(herring_collinearity[1:5], cex = 0.5, pch = 19, col = rgb(0, 0, 0, 0.5), 
                   labels = c("Chinook\nLength [mm]", "Chinook\nWeight [lbs]", "Collection\nDay in 2018", "Latitude", "Longitude"))

#Pairwise correlation coefficients for explanatory variables
print(cor(na.omit(herring_data[ ,c("sal.length", "sal.weight", "coll.doy", "lat", "long")])), digits = 2)

#Variance Inflation Factors
print(vif(na.omit(herring_data[ ,c("sal.length", "sal.weight", "coll.doy", "lat", "long")])))

#Balance of dataset (if appropriate)

```
Comment on pairwise scatterplots
Comment on strength of correlations
Comment on VIFs
Comment on potential difficulties from correlations and benefits

#Section 5: Statistical Methods and Model Fitting (200 words)
Statistical methods
- Which you will use and why
- Cite packages used for modelling
- What family of probability distributions (what link function)
- Which explantory variables are we including
- If and why data points are excluded
- If transformations are used and why
- If random-effects are used and why


```{r, echo = TRUE}
#Fit the model or models

```

#Section 6: Model Checking 

```{r, echo = TRUE}
#Check the assumptions of at least one model
#-  plotting and interpreting the residuals of the model
#-  check for correlation in the residuals due to grouping variables

```

Summary of the plausibility of the model
- Potential outliers

#Section 7: Model Summary, Conﬁdence Intervals and Model Comparison 

```{r, echo = TRUE}
#Model comparison and/or model selection and/or hypothesis testing
#- AIC (or AICc) values
#- AIC (or AICc) weights
#- R-squared (or adjusted R-squared) values
#- p-values. 
#Summary for at least one ﬁtted model
#- coeﬃcient estimates and their standard errors
#- conﬁdence intervals for coeﬃcient estimate
```

#Section 8: Plotting a Model with the Data

```{r, echo = TRUE}
#Plot at least one model with data
#- use method that shows strengths and weaknesses
#- include measurement of uncertainty
```

#Section 9: Discussion (400 words)
##A)
Assessment of how well the model seems to ﬁt the data
- model checking
- visualization of the model with the data
- 
##B)
Qualitative interpretation of the model
- interpretation of results in relation to your research questions and hypotheses
- general interpretation of conﬁdence intervals or standard errors for coeﬃcient estimates
-
##C)
Limitations of the model(s) that you have ﬁt
- Limitations
- Possible options for improvement
- Possible approaches for further data analysis

##D)
Results into context
- citie at least two relevant scientiﬁc papers